The Anatomy of a Temporal Summarization System

### 2015-02-28 conclusion.tex
almost final draft

Addtionally, we have shown that feature groups specifically targeted for the domain of disaster yield
better summaries; specifically features that capture the language and location
of the domain.

@@

We would also like to further analyze the failure 
of the temporal relevance features; intuitively we believe time to be important
for the update summarization task.

>>>>

%Addtionally, we have shown that feature groups specifically targeted for the domain of disaster yield
%better summaries; specifically features that capture the language and location
%of the domain.

We also have shown that features specifically targeted to the domain of disaster yield better summaries. We developed novel features that capture the language typical of different event types and that identify sentences specific to the particular disaster based on location.

@@

%We would also like to further analyze the failure 
%of the temporal relevance features; intuitively we believe time to be important
%for the update summarization task.

### 2015-02-28 dataset.tex
ck rev
almost final draft

For the document stream, we use the news portion of the
 2014 TREC KBA Stream Corpus.
The documents from this corpus come from hourly crawls of the web covering 
 October 2011 through February 2013.

Our experiments also make use of the TREC Temporal Summarization (TS) Track
 data from 2013 and 2014 \cite{aslam2013trec}. 
This data includes 25 events and event metadata (e.g., a user
search query for the event, the event type, and event evaluation timeframe).  
All events occurred during the time span of the TREC KBA Stream Corpus.
For each event we create a stream of relevant documents by 
selecting only documents with article content that contains the complete set 
of query words. 
%identifying the 
%article start/end within each web page in the corpus using a classifier
%tra
%those documents

>>>>

For the document stream, we use the news portion of the
 2014 TREC KBA Stream Corpus \cite{frank2012building}.
The documents from this corpus come from hourly crawls of the web covering 
 October 2011 through February 2013.

Our experiments also make use of the TREC Temporal Summarization (TS) Track
 data from 2013 and 2014 \cite{aslam2013trec}. 
This data includes 25 events and event metadata (e.g., a user
search query for the event, the event type, and event evaluation timeframe).  
All events occurred during the time span of the TREC KBA Stream Corpus.
For each event we create a stream of relevant documents by 
selecting only documents that contain the complete set 
of query words. 
%identifying the 
%article start/end within each web page in the corpus using a classifier
%tra
%those documents

### 2015-02-28 experiments.tex
almost final draft

We evaluate our system on two metrics: ROUGE \cite{lin2004rouge}, a common automatic
summarization method and an evaluation of system expected gain 
and comprehensiveness---metrics adapted 
from the TREC TS track \cite{aslam2013trec}.

@@

ROUGE measures the ngram overlap between a model (i.e. human) summary 
and an automatically generated system summary. 
Model summaries for each event were constructed by concatenating the event's 
nuggets. 
%See figure~\ref{fig:test} for an example model summary excerpt.
Generally, ROUGE evaluation assumes both model and system summaries
are of a bounded length. Since our systems are summarizing events over a span
of two weeks time, the total length of our system output is much longer than
the model. To address this, for each system/event pair, we sample with replacement
1000 random summaries of length less than or equal to the model summary 
(truncating the last sentence when neccessary). The final ROUGE score for the 
system are the average scores from these 1000 samples.

Because we are interested in system performance over time, we also evaluate 
systems at 12 hour intervals using the same regime as above. 
The model summaries in this case are retrospective, 
%and so we can see how quickly
%each system is able to achieve its maximum performance.
and this evaluation reveals how quickly each system is able to cover  information in the summary.

@@

%NIST developed a series metrics for evaluating update summarization systems
NIST developed a series of metrics for evaluating update summarization systems
as part of the TREC TS track.
We present results on two of these metrics, the expected gain and 
comprehensiveness.

@@

where $N$ is the set of event nuggets.
A comprehensiveness of 1 indicates that the summary has cover 
all nugget information for the event; the maximum
attainable comprehensiveness is 1.

Update-nugget matches are computed automatically, where a
match exists if the semantic similarity of the update/nugget pair is above 
a threshold.
Determining an optimal threshold to count matches is difficult so 
we evaluate at 
threshold values ranging
from .5 to 1, where values closer to 1 are more conservative estimates of
performance.
A manual inspection of matches 
suggests that semantic similarity values around .7 produce reasonable matches.
%For example, the nugget 
%\begin{blockquote}
%watches and warnings in effect for much of the Mid-Atlantic states and 
%southern New England 
%\end{blockquote}
%\noindent
%matches with the update 
%\begin{blockquote}
%A tropical storm warning remains in effect for much of the coasts 
%of North and South Carolina\end{blockquote}
%\noindent
%with a similarity of .7. 
The average semantic similarity of manual matches performed 
by NIST assessors was considerably lower at approximately .25, increasing
our confidence in the automatic matches in the .5--1 range.

\subsection{Model Comparisons}
We refer to our complete model as \textbf{AP+Salience}.  We compare this model again several variants and baselines intended to measure the contribution of different components. All thresholds
for all runs are tuned on the development set.

>>>>

We evaluate our system on two metrics: ROUGE \cite{lin2004rouge}, an automatic
summarization method and an evaluation of system expected gain 
and comprehensiveness---metrics adapted 
from the TREC TS track \cite{aslam2013trec}.

@@

ROUGE measures the ngram overlap between a model summary 
and an automatically generated system summary. 
Model summaries for each event were constructed by concatenating the event's 
nuggets. 
%See figure~\ref{fig:test} for an example model summary excerpt.
Generally, ROUGE evaluation assumes both model and system summaries
are of a bounded length. Since our systems are summarizing events over a span
of two weeks time, the total length of our system output is much longer than
the model. To address this, for each system/event pair, we sample with replacement
1000 random summaries of length less than or equal to the model summary 
(truncating the last sentence when neccessary). The final ROUGE scores for the 
system are the average scores from these 1000 samples.

Because we are interested in system performance over time, we also evaluate 
systems at 12 hour intervals using the same regime as above. 
The model summaries in this case are retrospective, 
%and so we can see how quickly
%each system is able to achieve its maximum performance.
and this evaluation reveals how quickly systems can cover  information in the model.

@@

%NIST developed a series metrics for evaluating update summarization systems
NIST developed metrics for evaluating update summarization systems
as part of the TREC TS track.
We present results on two of these metrics, the expected gain and 
comprehensiveness.

@@

where $N$ is the set of event nuggets.
A comprehensiveness of 1 indicates that the summary has covered
all nugget information for the event; the maximum
attainable comprehensiveness is 1.

Update-nugget matches are computed automatically; a
match exists if the semantic similarity of the update/nugget pair is above 
a threshold.
Determining an optimal threshold to count matches is difficult so 
we evaluate at 
threshold values ranging
from .5 to 1, where values closer to 1 are more conservative estimates of
performance.
A manual inspection of matches 
suggests that semantic similarity values around .7 produce reasonable matches.
%For example, the nugget 
%\begin{blockquote}
%watches and warnings in effect for much of the Mid-Atlantic states and 
%southern New England 
%\end{blockquote}
%\noindent
%matches with the update 
%\begin{blockquote}
%A tropical storm warning remains in effect for much of the coasts 
%of North and South Carolina\end{blockquote}
%\noindent
%with a similarity of .7. 
The average semantic similarity of manual matches performed 
by NIST assessors was much lower at approximately .25, increasing
our confidence in the automatic matches in the .5--1 range.

\subsection{Model Comparisons}
We refer to our complete model as \textbf{AP+Salience}.  We compare this model against several variants and baselines intended to measure the contribution of different components. All thresholds
for all runs are tuned on the development set.

### 2015-02-28 introduction.tex
almost final draft

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. 
We demonstrate that combining salience with clustering produces more relevant summaries
compared to baselines using clustering alone.  
Our experiments suggest that this
because our system is better able to adapt to
dynamic changes in input volume which adversely 
effect methods that use redundancy as a proxy for salience. 
%or predict salience but
%ignore redundancy in the input alone.

In addition to the tight integration between clustering and salience
prediction, our approach also exploits knowledge about the event to determine
salience. Thus, salience does not just represent importance within a set of
documents; it also represents both how typical a sentence is of the  
event
type (e.g., industrial accident, hurricane, riot) and whether it specifies 
information
about this particular event. 
Our feature representation includes a set of language models, one for each
event type, to measure the typicality of the sentence with regard to the 
current event, the distance of mentioned locations from the center of
the event, and the change in word frequencies over the time of the event.
While we evaluate these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
expected similarity of their ideal summaries.
%While we ground these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
%expected similarity of their ideal summaries.
% to represent its likelihood of referring to the input disaster. 
%We also utilize the change in word frequencies as we process the stream 
%to measure the changing importance of updates over time.

>>>>

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. 
We demonstrate that combining salience with clustering produces more relevant summaries
compared to baselines using clustering or relevance alone.  
Our experiments suggest that this
is because our system is better able to adapt to
dynamic changes in input volume which adversely 
affect methods that use redundancy as a proxy for salience. 
%or predict salience but
%ignore redundancy in the input alone.

In addition to the tight integration between clustering and salience
prediction, our approach also exploits knowledge about the event to determine
salience. Thus, salience represents both how typical a sentence is of the  
event
type (e.g., industrial accident, hurricane, riot) and whether it specifies 
information
about this particular event. 
Our feature representation includes a set of language models, one for each
event type, to measure the typicality of the sentence with regard to the 
current event, the distance of mentioned locations from the center of
the event, and the change in word frequencies over the time of the event.
While we evaluate these features in the domain of disaster, this approach is generally applicable to many update summarization tasks.
% where queries can be categorized by the 
%expected similarity of their ideal summaries.
%While we ground these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
%expected similarity of their ideal summaries.
% to represent its likelihood of referring to the input disaster. 
%We also utilize the change in word frequencies as we process the stream 
%to measure the changing importance of updates over time.

### 2015-02-28 methods.tex
almost final draft

Our update summarization system takes as input 
\begin{enumerate*}[label=\itshape\alph*\upshape)]
  \item a short query defining the event to be tracked (e.g. `Hurricane Sandy'), 
  \item an event category defining the type of event to be tracked (e.g. `hurricane'), 
  \item a stream of time-stamped documents %, $(\doc_0, \doc_1,\ldots,\doc_T)$,
  presented in temporal order, and \item an evaluation time period of interest.
    %  $(\stime,\etime)$.  
\end{enumerate*} 
While processing documents
throughout the time period of interest, the system outputs sentences
from these documents likely to be useful to query issuer.  We refer
to these selected sentences as \emph{updates}.

In order to measure the usefulness of a system's updates, we consider
the degree to which the system output reflects the different
aspects of an event.  Events are often composed of a variety of sub-events.  
For example, the Hurricane Sandy
event includes sub-events related to the storm making landfall,
the ensuing flooding, the many transportation issues, amongst many
others.  An ideal system would update the user about each of these
sub-events as they occur, with low latency.  
%In order to use
%language consistent with previous literature, 
We refer to these
sub-events as the \emph{nuggets} associated with an event.  A nugget is
defined as a fine-grained atomic sub-event associated with an event.  
%An example nugget for Hurricane Sandy might be ``Sandy on course for the Jersey Shore.''
We present 2 example nuggets associated with the Hurricane
Sandy event 
in Figure \ref{fig:nuggets}. Note that each event has anywhere from 50 to
several hunder nuggets in total.
We describe how these 
nuggets are found in Section \ref{sec:data}.

@@

\subsubsection{Features}
\label{sec:features}
We want our model to be predictive of sentence salience across different event instances so we avoid lexical features.  Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.

\paragraph{Basic Features}
We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}. These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.  
The data stream comprises text extracted from raw html documents;
these features help to downweight sentences that are not actually article 
content (e.g. web page titles, links to other content) or
more heavily weight important sentences (e.g., that appear in
prominent positions such as paragraph initial or article initial).

\paragraph{Query Features}

Query features measure the relationship between the sentence and the event query and type.  These include the number of query words present in the sentence in addition to the number of event type synonyms, hypernyms, and hyponyms as found in WordNet \cite{miller1995wordnet}.  
For example, for event type \emph{earthquake},  we match sentence terms 
``quake'', ``temblor'', ``seism'', and ``aftershock''.

\paragraph{Language Model Features}\label{subsubsec:lm}
Language models allow us to measure the likelihood of a sentence having been 
produced from a particular source.  We consider two types of language model 
features.  The first model is estimated from a corpus of generic news 
articles (we used the 1995-2010 Associated Press section of the 
Gigaword corpus \cite{graff2003english}).  
This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps our model to select sentences
written in the newswire style.

@@

We use the SRILM toolkit \cite{stolcke2002srilm} to implement a 5-gram Kneser-Ney model for both
the background language and model and the event specific language models.
For each sentence we use the average token log probability under each model
as a feature.

@@

\paragraph{Geographic Relevance Features}
%\fdcomment{need introduction here.  what are geographic features?}
The disasters in our corpus are all phenomena that
affect some part of the word. 
Where possible, we would like to capture a sentence's proximity to the event,
i.e. when a sentence references a location, it should be close to the 
area of the disaster.

@@

\paragraph{Temporal Relevance Features}
%\fdcomment{motivate burstiness; explain what it is before using the expression.}
As we track events over time, it is likely that the coverage of the event 
may die down, only to spike back up when there is a breaking development.
Alternatively, words that were salient initially may become background context,
and over time lose their importance.
Identifying terms that are ``bursty,'' i.e. suddenly peaking in usage,
can help to locate novel sentences that are part of the most recent reportage
and have yet to fall into the background.

@@

Exemplars that pass these thresholds are added selected as updates and added
to the summary.

>>>>

Our update summarization system takes as input 
\begin{enumerate*}[label=\itshape\alph*\upshape)]
  \item a short query defining the event to be tracked (e.g. `Hurricane Sandy'), 
  \item an event category defining the type of event to be tracked (e.g. `hurricane'), 
  \item a stream of time-stamped documents %, $(\doc_0, \doc_1,\ldots,\doc_T)$,
  presented in temporal order, and \item an evaluation time period of interest.
    %  $(\stime,\etime)$.  
\end{enumerate*} 
While processing documents
throughout the time period of interest, the system outputs sentences
from these documents likely to be useful to the query issuer.  We refer
to these selected sentences as \emph{updates}.

In order to measure the usefulness of a system's updates, we consider
the degree to which the system output reflects the different
aspects of an event.  Events are often composed of a variety of sub-events.  
For example, the Hurricane Sandy
event includes sub-events related to the storm making landfall,
the ensuing flooding, the many transportation issues, amongst many
others.  An ideal system would update the user about each of these
sub-events as they occur, with low latency.  
%In order to use
%language consistent with previous literature, 
We refer to these
sub-events as the \emph{nuggets} associated with an event.  A nugget is
defined as a fine-grained atomic sub-event associated with an event.  
%An example nugget for Hurricane Sandy might be ``Sandy on course for the Jersey Shore.''
We present 2 example nuggets associated with the Hurricane
Sandy event 
in Figure \ref{fig:nuggets}. Note that each event has anywhere from 50 to
several hundred nuggets in total.
We describe how these 
nuggets are found in Section \ref{sec:data}.

@@

\subsubsection{Features}
\label{sec:features}
We want our model to be predictive of sentence salience across different event instances so we avoid lexical features.  Instead, we extract features such as language model scores, geographic relevance, and temporal relevance from each sentence.

\textbf{Basic Features}
We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}. These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.  
The data stream comprises text extracted from raw html documents;
these features help to downweight sentences that are not 
content (e.g. web page titles, links to other content) or
more heavily weight important sentences (e.g., that appear in
prominent positions such as paragraph initial or article initial).

\textbf{Query Features}
Query features measure the relationship between the sentence and the event query and type.  These include the number of query words present in the sentence in addition to the number of event type synonyms, hypernyms, and hyponyms as found in WordNet \cite{miller1995wordnet}.  
For example, for event type \emph{earthquake},  we match sentence terms 
``quake'', ``temblor'', ``seism'', and ``aftershock''.

\textbf{Language Model Features}\label{subsubsec:lm}
Language models allow us to measure the likelihood of producing a sentence 
from a particular source.  We consider two types of language model 
features.  The first model is estimated from a corpus of generic news 
articles (we used the 1995-2010 Associated Press section of the 
Gigaword corpus \cite{graff2003english}).  
This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps our model to select sentences
written in the newswire style.

@@

We use the SRILM toolkit \cite{stolcke2002srilm} to implement a 5-gram Kneser-Ney model for both
the background language model and the event specific language models.
For each sentence we use the average token log probability under each model
as a feature.

@@

\textbf{Geographic Relevance Features}
%\fdcomment{need introduction here.  what are geographic features?}
The disasters in our corpus are all phenomena that
affect some part of the word. 
Where possible, we would like to capture a sentence's proximity to the event,
i.e. when a sentence references a location, it should be close to the 
area of the disaster.

@@

\textbf{Temporal Relevance Features}
%\fdcomment{motivate burstiness; explain what it is before using the expression.}
As we track events over time, it is likely that the coverage of the event 
may die down, only to spike back up when there is a breaking development.
%Alternatively, words that were salient initially may become background context,
%and over time lose their importance.
Identifying terms that are ``bursty,'' i.e. suddenly peaking in usage,
can help to locate novel sentences that are part of the most recent reportage
and have yet to fall into the background.

@@

Exemplars that pass these thresholds are selected as updates and added
to the summary.

### 2015-02-28 relatedwork.tex
almost final draft

The last category includes probabilistic \cite{haghighi2009exploring}, 
information theoritic, and set cover \cite{lin2011class}
approaches. While these methods are generally focused on producing diverse
summaries, they are difficult to adapt to the streaming setting, where 
we do not necessarily have a fixed summary length and the corpus to be
summarized contains many irrelevent sentences, i.e. there are large
portions of the corpora that we specificly want to avoid.

@@

%\fdadd{
Several researchers have recognized the importance of summarization during
natural disasters.  \cite{qi:temporal-summarization} developed a system for detecting
novel, relevant, and comprehensive sentences immediately after a natural
disaster.  
%The method uses a model of
%sentence relevance and novelty in order to select appropriate updates.
%Training data for regression targets is automatically generated from
%retrospective Wikipedia data.  The system is evaluated on news documents
%related to 197 natural and human disasters from 2009 to 2011 using variants of
%ROUGE modified to capture novelty, relevance, and comprehensiveness
%\cite{lin2004rouge}.  
\cite{wang:update-summarization} present a clustering-based approach to
efficiency detect important updates during natural disasters.  
The algorithm works by hierarchically
clustering sentences online, allowing the system to output a more expressive
narrative structure than \cite{qi:temporal-summarization}.  
Our system attempts to unify these system's approaches 
(predictive ranking and clustering respectively). 
%The method is evaluated on
%official press releases related to Hurricane Wilma  in 2005 using ROUGE score
%between the system summary and a manually generated target summary. 
%In addition
%to natural disasters, we evaluate our system on a variety of event types 
% such as terrorism and mass shootings (e.g., the 2012
%shooting in Aurora, Col.), accidents
%(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
%Arab spring). The size and scope of the events varies considerably. Hurricane
%Sandy, for example, affected multiple countries over weeks while the
%2012 Aurora shooting was contained to a single location spanning several hours.

>>>>

The last category includes probabilistic \cite{haghighi2009exploring}, 
information theoritic, and set cover \cite{lin2011class}
approaches. While these methods are  focused on producing diverse
summaries, they are difficult to adapt to the streaming setting, where 
we do not necessarily have a fixed summary length and the corpus to be
summarized contains many irrelevent sentences, i.e. there are large
portions of the corpora that we specificly want to avoid.

@@

%\fdadd{
Several researchers have recognized the importance of summarization during
natural disasters.  \cite{qi:temporal-summarization} developed a system for detecting
novel, relevant, and comprehensive sentences immediately after a natural
disaster.  
%The method uses a model of
%sentence relevance and novelty in order to select appropriate updates.
%Training data for regression targets is automatically generated from
%retrospective Wikipedia data.  The system is evaluated on news documents
%related to 197 natural and human disasters from 2009 to 2011 using variants of
%ROUGE modified to capture novelty, relevance, and comprehensiveness
%\cite{lin2004rouge}.  
\cite{wang:update-summarization} present a clustering-based approach to
efficiently detect important updates during natural disasters.  
The algorithm works by hierarchically
clustering sentences online, allowing the system to output a more expressive
narrative structure than \cite{qi:temporal-summarization}.  
Our system attempts to unify these system's approaches 
(predictive ranking and clustering respectively). 
%The method is evaluated on
%official press releases related to Hurricane Wilma  in 2005 using ROUGE score
%between the system summary and a manually generated target summary. 
%In addition
%to natural disasters, we evaluate our system on a variety of event types 
% such as terrorism and mass shootings (e.g., the 2012
%shooting in Aurora, Col.), accidents
%(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
%Arab spring). The size and scope of the events varies considerably. Hurricane
%Sandy, for example, affected multiple countries over weeks while the
%2012 Aurora shooting was contained to a single location spanning several hours.
